---
title: "Gender representation analysis"
author: "Astrid Rybner"
date: "5/20/2022"
output: html_document
---

```{r}
library(pacman)
p_load(fpp3,fable, tidyverse, dplyr, tidyr, readxl)

```

# Calculate mean values
Reading in the data, grouping by year and taking the mean and creating a timeseries object. 
```{r}
data <- read.csv("gender_counts_preprocessed.csv") %>% select(publish_date, gender, male_skewness_log)


data <- data %>% filter(publish_date < 1961)

grouped <- aggregate(data[,3], list(data$publish_date, data$gender), mean) 
colnames(grouped) <- c('Year', 'Gender', 'value')

timeseries <- grouped %>%
  as_tsibble(index = Year, 
             key = Gender)
```

# Identify missing variables
```{r}
# 141 distinct years in the period 1820-1960 (incuding both of these years)
timeseries %>% filter(Gender=='F') %>% distinct(Year) # 124 for female
timeseries %>% filter(Gender=='M') %>% distinct(Year) # 141 for male
# so we have some missing values in some years for females!
```

# Fill in missing values
Filling in the missing years with NA's, and then adding the value from the previous year
```{r}
timeseries_f <- timeseries %>% group_by_key() %>% tsibble::fill_gaps(Year) %>%
  tidyr::fill(value, .direction = "down")

```

# Plotting the data 
```{r}
timeseries_f %>%
  autoplot() +
  labs(x = "Year", y = "Percentage", title = "Percentage of pronouns that are male")
```
The signal looks approximately stationary, and the variance looks approximately stationary. 

# Is there a trend in the signal?
```{r}
f <- timeseries_f %>% filter(Gender =='F')
m <- timeseries_f %>% filter(Gender =='M')


library(Kendall)
MannKendall(f$value) # significant negative trend for females
MannKendall(m$value) # non significant negative trend for females

```
https://www.statisticshowto.com/mann-kendall-trend-test/ 

The test was suggested by Mann (1945) and has been extensively used with environmental time
series (Hipel and McLeod, 2005). 
https://cran.r-project.org/web/packages/Kendall/Kendall.pdf 



# Doing a STL decomposition (to understand the time series better)
NB this is an additive decomposition method. 
Looks like there is no seasonality and for female data; slightly increasing and for males a slightly decreasing trend. 

```{r}
# decomposition
dcmp <- timeseries_f %>%
  model(stl = STL())

# looking at the trend and the remainder component separately 
dcmp %>%
  components() %>%
  autoplot() + theme_bw() + scale_color_brewer(palette="Dark2")

ggsave("../output/components.png", width = 30, height = 20, units = "cm")
```

# Plotting the raw data with the trend
```{r}
library(RColorBrewer)
timeseries_f %>%
  autoplot(value) + scale_color_brewer(direction = -1, palette="Set2")+
  autolayer(components(dcmp), trend, color = "black") + geom_vline(xintercept = 1945, linetype="dashed", 
                color = "blue", size=0.7) + theme_bw() + xlab("Year") + ylab(" ") + ggtitle("Trend in skewness by gender of the author") +
  theme(plot.title = element_text(size=25),
      axis.title.y = element_text(size=18),
      axis.title.x = element_text(size=18),
      axis.text.x = element_text(size = 15),
      axis.text.y = element_text(size = 18),
      legend.text = element_text(size = 15),
      legend.title = element_text(size = 18))+
  labs(x = "Year", y = "Proportion (log odds)")

ggsave("../output/trend-raw-data.png", width = 30, height = 20, units = "cm")

```
